Sequent calculus style in Dedukti

### 2023-07-27 draft.tex
tex

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

@@

	\subsection{The spirit of Dedukti encodings}

@@

	\paragraph{First-order logic.} First, we In order to define FOL in Dedukti.

	The idea is the symbols \textsf{Term} and \textsf{Formula} are types as their type is TYPE. Moreover, these symbols are declared as constant to avoid rewriting.

	First, we define the type of terms
	This type can be extended

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={8-8}]{SimpleVersion/FOL.lp}

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={4-4}]{SimpleVersion/FOL.lp}

@@

	\subsection{Sequent calculus}

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{hyperref}

@@

	\subsection{Dedukti}

@@

@@

	\subsection{First Order Logic with Equality and Schematic Variables}
	We consider first order logic with logical operators $(\neg, \land, \lor, \rightarrow, \leftrightarrow, \forall, \exists, \exists !)$, where $\exists !$ denotes the unique existential. We define the types of \textit{terms} and \textit{formulas}:

	\begin{lstlisting}[language=Dialekto]
constant symbol Term : TYPE;
constant symbol Formula : TYPE;
	\end{lstlisting}
	Our encoding imposes no requirements on the structure of terms. In practice,  for every function symbol $f$ of arity $n$ in the source language, the user needs to define a corresponding symbol:
	\begin{lstlisting}[language=Dialekto]
constant symbol f :  Term → ... → Term
	\end{lstlisting}
	and similarly for predicate symbols
	\begin{lstlisting}[language=Dialekto]
constant symbol p :  Term → ... → Formula
	\end{lstlisting}
	As we consider first order logic with equality and have special rules for it, we define the equality symbol:
	\begin{lstlisting}[language=Dialekto]
constant symbol Equality :  Term → Term → Formula
	\end{lstlisting}
	And logical operators:
	\begin{lstlisting}[language=Dialekto]
symbol Neg : Formula → Formula;
symbol ∧ : Formula → Formula → Formula; notation ∧ infix right 30;
symbol ∨ : Formula → Formula → Formula; notation ∨ infix right 30;
symbol Implies : Formula → Formula → Formula;
symbol Iff : Formula → Formula → Formula;
\end{lstlisting}
%symbol Forall : (Term → Formula) → Formula;
%symbol Exists : (Term → Formula) → Formula;
%symbol ExistsOne : (Term → Formula) → Formula;
%	\end{lstlisting}
	\subsection{Higher Order Abstract Syntax}
	Higher Order Abstract Syntax (HOAS) is an approach to make shallow encoding of a system (in our case FOL) in an other ($\lambda \Pi$) where variables of the original are represented as variable of the target system and not as syntactic objects.

	\iffalse
	We consider first order logic with arbitrary constant function and predicate symbols. Fomally We fix:
	\begin{itemize}
		\item a set of constant symbols $C$ noted $f^j, g^j, h^j, f^j_i,...$ for indices $i\in \mathbb{N}$ and where the superscript $j$ denotes the \textit{arity} of the symbol
		\item a countably infinite set of schematic function symbols $'C$ noted $ 'f^j, 'g^j, 'h^j, f^j_i,...$ for indices $i\in \mathbb{N}$ and where the superscript $j$ denotes the \textit{arity} of the symbol.
	\end{itemize}
	We also call schematic functions symbols of arity 0 \textit{variables}, and note them $x,y,z$.
	The set of terms of height at most $n$ is
	$$\mathcal{T}_0 := \lbrace f^0 \mid f^0 \in C\rbrace \cup \lbrace 'f^0 \mid 'f^0 \in C \rbrace$$
	$$\mathcal{T}_{n+1} := \mathcal{T}_n \cup \lbrace f^j(\vec{t}) \mid f^j\in \mathcal{C}, \vec{t}\in \mathcal{T}_n^j, j>0 \rbrace \cup \lbrace 'f^j(\vec{t}) \mid 'f^j\in \mathcal{C}, \vec{t}\in \mathcal{T}_n^j, j>0 \rbrace$$
	and the set of all termsis $\mathcal T = \bigcup_{n=0}^\infty \mathcal{T}_n$.  Moreover, we allow \textit{schematic functions, predicate and connector symbols}.

	\fi
	\subsection{Sequent calculus}
	The Sequent Calculus LK is a prof system for first order logic. A \textit{sequent} is a pair of ordered lists of formulas, noted
	$$
	\Gamma \vdash \Delta
	$$
	We usually use capital greek letter to denote sets of formulas. In an expression such as $\Gamma, \Sigma \vdash \phi, \Delta$, the comma denotes concatenation.
	The \textit{proof steps} of Sequent Calculus are depicted in \autoref{fig:scsteps1}, \autoref{fig:scsteps2} and \autoref{fig:scsteps3}. There are other formalizations of Sequent Calculus, including some where $\Gamma$ and $\Delta$ are \textit{sets} rather than list,as justified by the \texttt{Contract} and \texttt{Swap} rules, but this formalism is best suited for encoding in Dedukti and is the one we assume in this paper.

	The rules in \autoref{fig:scsteps1} are called structural rules. The rules in \autoref{fig:scsteps2} are called introduction rules. Each of them correspond to the introduction of s logical symbol on the left or right hand side of the sequent.
	The rules in \autoref{fig:scsteps3} are called \textit{bonus rules}. These are not typically presented as part of sequent calculus. Since our language is first order logic \textit{with equlity}, the two \texttt{Equality} rules are necessary to encode the properties of equality. the two \texttt{IffSubst} rules describe the substitution of identical formulas for identical formulas. This is an admissible rule in first order logic without schematic symbols, as $\iff$ is a congruence relation for logical operators. In presence of schematic connector symbols, this rule is sound but not admissible. We detail the question more in \autoref{sect:propext}, when we describe the axiom of propositional extensionality.
